---
title: "Chapter 3: The Beta-Binomial Bayesian Model"
author: "Thomas Sandmann"
format:
  html:
    embed-resources: true
editor_options: 
  chunk_output_type: console
execute: 
  cache: false
---

This quarto markdown documents contains code from Chapter 3 of
[Bayes Rules! An Introduction to Applied Bayesian Modeling](https://www.bayesrulesbook.com/chapter-2#building-a-bayesian-model-for-events)
by Alicia A. Johnson, Miles Q. Ott and Mine Dogucu.

I am using it to document my progress and to take notes as I work my way through
the book. Many thanks to the authors for making their book available online!

```{r}
#| message: false
#| warning: false
library(bayesrules)
library(janitor)
library(patchwork)
library(sessioninfo)
library(tidyverse)
```

```{r}
# Plot the Beta(45, 55) prior
plot_beta(45, 55)
```

```{r}
plot_beta_binomial(alpha = 45, beta = 55, y = 30, n = 50)
```

The summarize_beta_binomial() function in the bayesrules package summarizes the typical values and variability in the prior and posterior models of π.

```{r}
summarize_beta_binomial(alpha = 45, beta = 55, y = 30, n = 50)
```

```{r}
set.seed(84735)
michelle_sim <- data.frame(pi = rbeta(10000, 45, 55)) %>% 
  mutate(y = rbinom(10000, size = 50, prob = pi))
ggplot(michelle_sim, aes(x = pi, y = y)) + 
  geom_point(aes(color = (y == 30)), size = 0.1)
```

```{r}
# Keep only the simulated pairs that match our data
michelle_posterior <- michelle_sim %>% 
  filter(y == 30)

# Plot the remaining pi values
ggplot(michelle_posterior, aes(x = pi)) + 
  geom_density()
```

```{r}
michelle_posterior %>% 
  summarize(mean(pi), sd(pi))
```

## Reproducibility

<details>
<summary>
Session information
</summary>

```{r}
sessionInfo()
```

</details>